An FMI-Based Initialization Plugin for INTO-CPS Maestro 2

### 2020-06-30 background.tex
Updated images in background section

The interconnections of FMU variables can lead to circular dependencies between these. The following definitions will formalize the concept of an algebraic loop in a co-simulation scenario, and define the problem these loops are introducing in the system. A co-simulation scenario containing an algebraic loop is shown in \cref{fig:fmu_cycle}.

\begin{figure}
    \centering
    \includegraphics{images/fmu_cycle.pdf}
    \caption{Co-simulation scenario with cycle}
    \label{fig:fmu_cycle}
\end{figure}

\begin{definition}[Algebraic loops] 
The set of interconnected variables in an algebraic loop is defined as a strongly connected component of the graph in \cref{def:initialization_graph}. \\
A strongly connected component: $SCC = v \in V: Path(v,v)$\\
$\LoopVariables = \{v | v \in V : Path(v,v)\}$ \\
An algebraic loop is defined as the set $\{p_1, p_2 | p_1, p_2 \in \LoopVariables: Path(p_1, p_2) \land Path(p_2, p_1)\}$\todo{Look at this - can it be specified better?}\\
Path is defined as the transitive closure of the edges of the graph denoted by \cref{def:initialization_graph}.\\
\todo{Claudio check this definition and add a formalization of the challenge introduced}
\end{definition}

>>>>

The interconnections of FMU variables can lead to circular dependencies between these as seen in \cref{fig:fmu_cycle} showing a co-simulation scenario containing an algebraic loop and its Initialization Graph.

\begin{figure}[H]
    \centering
    \begin{minipage}{0.55\textwidth}
        \centering
    \includegraphics[width=1\textwidth]{images/fmu_cycle.pdf}
    \end{minipage}\hfill
    \begin{minipage}{0.35\textwidth}
        \centering
    \includegraphics[width=1\textwidth]{images/SCC.pdf}
    \end{minipage}
    \caption{An FMU co-simulation scenario and its Initialization Graph}
    \label{fig:fmu_cycle}
\end{figure}

The following definitions will formalize the concept of an algebraic loop in a co-simulation scenario, and define the problem these algebraic loops are introducing in the system.

\begin{definition}[Algebraic loops] 
An algebraic loop is defined as a strongly connected component of the graph in \cref{def:initialization_graph}. \\
A strongly connected component: $SCC = v \in V: Path(v,v)$\\
$\LoopVariables = \{v | v \in V : Path(v,v)\}$ \\
An algebraic loop is defined as the set $\{p_1, p_2 | p_1, p_2 \in \LoopVariables: Path(p_1, p_2) \land Path(p_2, p_1)\}$\todo{Look at this - can it be specified better?}\\
Path is defined as the transitive closure of the edges of the graph denoted by \cref{def:initialization_graph}.\\
\todo{Claudio check this definition and add a formalization of the challenge introduced}
\end{definition}

### 2020-06-30 initializationorder.tex
Updated images in background section

In the \textit{Initialization Mode} state all variables of all FMUs should be defined.
The variables with \textit{causality = parameter} of each FMU are set first, and the order of which they are set is again insignificant as they are independent.
Afterwards the interconnected variables should be defined, but as stated by the definitions \ref{def:feedthrough}, \ref{def:getout} and \ref{def:setin} the operations \textit{get} and \textit{set} \textbf{require} a specific initialization order, and algebraic loops places even more constraints on the initialization order. This phase should should also support algebraic loops, this support includes utilization of fixed point iteration and detection of non-convergence of algebraic loops.

@@

\subsubsection{Structure of the graph}
The initialization algorithm starts by building a directed graph of the dependency between the interconnected variables of the FMUs. The graph is constructed based on the interconnected variables and internal connections (feed-through). Each interconnected variable in the system represents a node, and the edges are based on these connections. The edges of this graph represent precedence constraints based on the algebraic dependencies of the interconnected variables. Please see definition \ref{def:initialization_graph} for a formal definition of the graph.

As described earlier, not all co-simulation scenarios are valid, and these invalid scenarios need to be identified. This is done by monitoring convergence of all algebraic loops. A valid co-simulation scenario must convergence by definition given in definition \ref{def:convergence}.

\subsubsection{Calculation of an initialization order}
The topological ordering of the graph defined in definition \ref{def:initialization_graph} is used as the initialization order of the interconnected FMU variables. The elements of the topological order are the strongly connected components of the graphs. The strongly connected components are potentially algebraic loops of the system. The algebraic loops are easily identified as the strongly connected components containing more than a single element. The single element members of the topological order are standard interconnected FMU variables, whose port operation should only be performed once during the initialization procedure. \\
The calculation of an initialization order is performed in linear time based on the number of both external and internal connections using Tarjan's algorithm \cite{tarjan_1972}. This algorithm was selected since it solves two of the central issues of the initialization-phase of the co-simulation.
\begin{itemize}
    \item Identifies algebraic loops between interconnected variables (strongly connected components)
    \item Performs a topological sorting of the Strongly Connected Components
\end{itemize}
The algorithm is furthermore well-established, and there exist formal proofs of its correctness and properties\cite{stefanMerz}. The algorithm is among the most efficient graph algorithms for achieving the defined goals.  
Tarjan's algorithm is performing a topological sorting of the strongly connected components of a graph. This makes it able to handle both graphs with and without algebraic loops.

\todo{Come with an example using a graph with loops}

@@

\subsection{Optimization of a Initialization Procedure}
An initialization procedure can be optimized by using the fact that the FMI specification allows multiple \textit{set} or \textit{get} operations of the same FMU to be performed in bulk by grouping them together to a single operation over multiple variables with similar characteristic. The criteria of optimization can of an initialization procedure is formalized in definition \ref{def:optimization}
\begin{definition}[Optimization of a Initialization procedure]\label{def:optimization}
  Given an initialization procedure $\sequence{\initcall_i}_{i \in \setnat}$ with a finite ordered sequence of FMU function calls $\functioncall_i \in \allfunctioncalls = \bigcup_{c \in \fmus} \set{\fset{c},\fget{c}},$ and $i$ denoting the order of the function call. It can be optimized if $\exists \functioncall_i, \functioncall_{i +1} \in \allfunctioncalls : \exists c \in \fmus :(\functioncall_i \in {\fset{c}} \land \functioncall_{i+1} \in {\fset{c}}) \lor (\functioncall_i \in {\fget{c}} \land \functioncall_{i+1} \in {\fget{c}})$
\end{definition}
Since an Initialization procedure is defined in the same way as other co-simulation steps, the optimization criteria described in definition \ref{def:optimization} is valid for an arbitrary co-simulation step. \\
The correctness of this optimization can be established by the proof of using the Initialization Graph's topological ordering as the initialization order described by Gomes et al. \cite{Gomes2019}. This proof is still valid since this optimization does not change the structure of the Initialization Graph. \\
A limitation of this optimization strategy is that it is not guaranteed to find all potentially valid optimizations of a co-simulation scenario. Considering it works only on a specific co-simulation step (a topological order of a graph), which is not necessarily unique for a given co-simulation scenario/graph. To perform all viable optimizations of a co-simulation step, a more advanced optimization strategy should be used, or this optimization strategy should be performed on the set of all valid co-simulation steps - yielding a potential very inefficient initialization algorithm.

>>>>

In the \textit{Initialization Mode} state all variables of all FMUs should be defined.
The variables with \textit{causality = parameter} of each FMU are set first, and the order of which they are set is again insignificant as they are independent.
Afterwards the interconnected variables should be defined, but as stated by the \cref{def:feedthrough,def:getout,def:setin} the operations \textit{get} and \textit{set} \textbf{require} a specific initialization order, and algebraic loops places even more constraints on the initialization order. This phase should should also support algebraic loops, this support includes utilization of fixed point iteration and detection of non-convergence of algebraic loops.

@@

\subsubsection{Structure of the graph}
The initialization algorithm starts by building a directed graph of the dependency between the interconnected variables of the FMUs. The graph is constructed based on the interconnected variables and internal connections (feed-through). Each interconnected variable in the system represents a node, and the edges are based on these connections. The edges of this graph represent precedence constraints based on the algebraic dependencies of the interconnected variables. Please see \cref{def:initialization_graph} for a formal definition of the graph.

As described earlier, not all co-simulation scenarios are valid, and these invalid scenarios need to be identified. This is done by monitoring convergence of all algebraic loops. A valid co-simulation scenario must convergence by \cref{def:convergence}.

\subsubsection{Calculation of an initialization order}
The topological ordering of the graph defined in \cref{def:initialization_graph} is the initialization order of the interconnected FMU variables. The elements of the topological order are the strongly connected components of the graphs. The strongly connected components are potentially algebraic loops of the system. The algebraic loops are easily identified as the strongly connected components containing more than a single element. The single element members of the topological order are standard interconnected FMU variables, whose port operation should only be performed once during the initialization procedure. \\
The calculation of an initialization order is performed in linear time based on the number of both external and internal connections using Tarjan's algorithm \cite{tarjan_1972}. This algorithm was selected since it solves two of the central issues of the initialization-phase of the co-simulation.
\begin{itemize}
    \item Identifies algebraic loops between interconnected variables (strongly connected components)
    \item Performs a topological sorting of the Strongly Connected Components
\end{itemize}
The algorithm is furthermore well-established, and there exist formal proofs of its correctness and properties\cite{stefanMerz}. The algorithm is among the most efficient graph algorithms for achieving the defined goals.  
Tarjan's algorithm is performing a topological sorting of the strongly connected components of a graph. It is able to handle both graphs with and without algebraic loops.

@@

\subsection{Optimization of a Initialization Procedure}
An initialization procedure can be optimized since the FMI specification allows multiple \textit{set} or \textit{get} operations of the same FMU to be performed in bulk by grouping them together to a single operation over multiple variables with similar characteristic. The criteria of optimization can of an initialization procedure is formalized in \cref{def:optimization}
\begin{definition}[Optimization of a Initialization procedure]\label{def:optimization}
  Given an initialization procedure $\sequence{\initcall_i}_{i \in \setnat}$ with a finite ordered sequence of FMU function calls $\functioncall_i \in \allfunctioncalls = \bigcup_{c \in \fmus} \set{\fset{c},\fget{c}},$ and $i$ denoting the order of the function call. It can be optimized if $\exists \functioncall_i, \functioncall_{i +1} \in \allfunctioncalls : \exists c \in \fmus :(\functioncall_i \in {\fset{c}} \land \functioncall_{i+1} \in {\fset{c}}) \lor (\functioncall_i \in {\fget{c}} \land \functioncall_{i+1} \in {\fget{c}})$
\end{definition}
Since an Initialization procedure is defined in the same way as other co-simulation steps, the optimization criteria described in definition \ref{def:optimization} is valid for an arbitrary co-simulation step. \\
The correctness of this optimization can be established by the proof of using the Initialization Graph's topological ordering as the initialization order described by Gomes et al. \cite{Gomes2019}. This proof is still valid since this optimization does not change the structure of the Initialization Graph. \\
A limitation of this optimization strategy is that it is not guaranteed to find all potentially valid optimizations of a co-simulation scenario. Considering it works only on a specific co-simulation step (a topological order of a graph), which is not necessarily unique for a given co-simulation scenario/graph. To perform all viable optimizations of a co-simulation step, a more advanced optimization strategy should be used, or this optimization strategy should be performed on the set of all valid co-simulation steps - yielding a potential very inefficient initialization algorithm.
